---
title: "p8105_hw3_yl4610"
output: github_document
author: "Yuxin Liu"
date: "2022-10-08"
---

```{r, message=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# question 1 
```{r}
data("instacart") 
janitor::clean_names(instacart)
```

I loaded the libraries and data from the p8105.datasets. Then I used janitor to clean the names. This dataset contains variables such as `r names(instacart)`. It contains `r ncol(instacart)` columns and `r nrow(instacart)` rows, with each row representing a single product from an instacart order. 

Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```
I used count to look at the number of aisles and then used arrange to arrange the number from largest to smallest.

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
I used filter to limit this to aisles with more than 10000 items ordered.Then I used mutate and fct_reorder to arrange aisles sensibly and used ggplot to create the plot that shows the number of items ordered in each aisle.

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```
I used filter to filter three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. I used count to include the number of times each item is ordered and used mutate to rank them in order. I used arrange to rank the value from highest to lowest.I used knitr::kable to create a table in rmd file.

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```
I used filter to filter Pink Lady Apples and Coffee Ice Cream. I used group_by to group them by product_name and order_dow. Then I used summarize to calculate mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. I used spread to format the table for human readers. Lastly, I used knitr::kable to create a table in rmd file.

# question 2
```{r}
accel_data = read_csv( "./dataset/accel_data.csv") %>%
janitor::clean_names() %>% 
mutate (
  weekday_vs_weekend = ifelse(day %in% c("Saturday", "Sunday"), 0, 1),
  weekday_vs_weekend = as.factor (weekday_vs_weekend)
  ) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    names_prefix = "activity_", 
    values_to = "activity_counts") 
accel_data 
```
I used read_csv to read dataset and used janitor to clean the names of data. Then I used mutate to create the new variable weekday_vs_weekend by using ifelse. If day equals to Saturday or Sunday, weekday_vs_weekend equals to 0; If day equals to weekdays, weekday_vs_weekend equals to 1. Then I used mutate to convert them to factors. I used pivot_longer to pivot the categories of activities into one column and pivot corresponding values to activity_counts. I also used names_to to dropped the prefix activity_ to make it cleaner.

This dataset contains variables such as `r names(accel_data)`. It contains `r ncol(accel_data)` columns and `r nrow(accel_data)` rows. Thus, there are `r nrow(accel_data)` observations. 

```{r}
accel_data %>%
  group_by (week, day) %>%
  summarise (
  activity_total = sum(activity_counts)) %>%
  pivot_wider (
    names_from = "day",
    values_from = "activity_total")%>% 
  select (Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday) %>% 
  knitr::kable() 
```
I used group_by to make groupings explicit so that they can be included in subsequent operations. Then I used summarise to create a total activity variable for each day by aggregating across minutes via sum function. Then, I used pivot_wider in order to better see the data and used select to reorder from Monday to Sunday. I used knitr::kable to create a table in rmd file.

This person had extreme low level of activities on Saturday at week 4 and week 5. Other than this, I do not think there are any apparent trends. 

```{r}
accel_data %>%
  ggplot(aes(x = activity, y = activity_counts, color = day)) + 
  geom_line(alpha = 0.5) + 
    labs( 
    title = "24-hour activity time courses for each day",
    x = "minute",
    y = "day") +
    geom_smooth(alpha = 0.5)
```
Based on this graph, this 63 year-old male had a higher level of activities in weekdays compared to weekends.

# question3
```{r}
data("ny_noaa") 
ny_noaadf = 
ny_noaa %>%
janitor::clean_names() %>% 
  mutate (tmax = as.integer(tmax),
          tmin = as.integer(tmin)) %>% 
separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate (
      year = as.integer(year),
      month = as.integer(month),
      day = as.integer(day)) %>% 
  mutate (
    prcp = prcp/10,
      tmax = tmax/10,
        tmin = tmin/10)
```
I loaded ny_noaa dataset from the p8105.datasets library.
This dataset contains variables such as `r names(ny_noaa)`. It contains `r ncol(ny_noaa)` columns and `r nrow(ny_noaa)` rows. Thus, there are `r nrow(ny_noaa)` observations. 

prec: Precipitation has `r sum(is.na(ny_noaadf$prcp))` missing values
snow: Snowfall has `r sum(is.na(ny_noaadf$snow))` missing values
snwd: Snow depth variable has `r sum(is.na(ny_noaadf$snwd))` missing values
tmin: Minimum temperature has `r sum(is.na(ny_noaadf$tmin))` missing values
tmax: Maximum temperature has `r sum(is.na(ny_noaadf$tmax))` missing values
  
Missing data is a big issue here because we have a lot of them. 

```{r}
ny_noaadf %>% 
  group_by(snow) %>% 
    summarize(
    n_obs = n()) %>% 
arrange(desc(n_obs))
```
The most commonly observed values for snow is 0 because it has 2008508 observations.

Two-panel plot showing the average max temperature in January and in July in each station across years:
```{r}
ny_plot1 = 
ny_noaadf %>%
mutate(month = month.name[month]) %>% 
filter(month %in% c("January","July")) %>% 
group_by(id, year, month) %>% 
summarize(
average_max_temp = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(
  aes(x = year, y = average_max_temp, color=month, group=id)) +
  geom_line(alpha = 0.5, size = .1) +
  facet_grid(.~month)
ny_plot1
```
Is there any observable / interpretable structure? Any outliers?
The average max temperature in January is lower than the average max temperature in July in each station across year. Also, the range of average max temperature in January is greater than that in July.

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
```{r}
ny_ploti =
  ny_noaadf %>%
  ggplot(
  aes(x = tmin, y = tmax)) +
  geom_line(alpha = 0.5, size = .1)
ny_ploti
```

```{r}
ny_plotii =
  ny_noaadf %>%
  filter(snow>0 & snow<100) %>% 
  ggplot(aes(x = snow, fill = year)) +
  geom_density(alpha = .4, adjust = .5, color = "blue")
ny_plotii
```



